%\documentclass[letterpaper,10pt]{article}
\documentclass[onecolumn]{emulateapj}





\usepackage{geometry}   % See geometry.pdf to learn the layout
                        % options.  There are lots.
\usepackage[latin1]{inputenc}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{soul}

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}




%\author{Michael L. Norman\altaffilmark{1,3}}
%\author{Daniel R. Reynolds\altaffilmark{4}}
%\author{Geoffrey C. So\altaffilmark{1}}
%\author{Robert P. Harkness\altaffilmark{2,3}}

%\affiliation{
%\altaffilmark{1}CASS, University of California, San Diego, 9500 Gilman Drive La Jolla, CA 92093-0424\\
%\altaffilmark{2}NICS, Oak Ridge National Laboratory, 1 Bethel Valley Rd, Oak Ridge, TN 37831\\
%\altaffilmark{3}SDSC, University of California, San Diego, 9500 Gilman Drive La Jolla, CA 92093-0505\\ 
%\altaffilmark{4}Southern Methodist University, 6425 Boaz Ln, Dallas, TX 75205\\
%}

\renewcommand{\(}{\left(}
\renewcommand{\)}{\right)}
\newcommand{\vb}{{\bf v}_b}
\newcommand{\xvec}{{\bf x}}
\newcommand{\Omegabar}{\bar{\Omega}}
\newcommand{\rhob}{\rho_b}
\newcommand{\dt}{\Delta t}
\newcommand{\Eot}{E^{OT}}
\newcommand{\Enu}{E_{\nu}}
\newcommand{\Fnu}{{\bf F}_{\nu}}

\newcommand{\Pnu}{\overline{\bf P}_{\nu}}
\newcommand{\Ef}{E_f}
\newcommand{\sighat}{\hat{\sigma}}
\newcommand{\R}{I\!\!R}
\newcommand{\Rthree}{\R^3}
\newcommand{\eh}{e_h}
\newcommand{\ec}{e_c}
\newcommand{\Edd}{\mathcal F}
\newcommand{\Eddnu}{\Edd_{\nu}}
\newcommand{\mn}{{\tt n}}
\newcommand{\mB}{\mathcal B}
\newcommand{\mC}{{\mathcal C}}
\newcommand{\mL}{{\mathcal L}}
\newcommand{\mD}{{\mathcal D}}
\newcommand{\mDnu}{\mD_{\nu}}
\newcommand{\mCnu}{\mC_{\nu}}
\newcommand{\mLnu}{{\mathcal L}_{\nu}}
\newcommand{\mCe}{\mC_e}
\newcommand{\mLe}{\mL_e}
\newcommand{\mCn}{\mC_{\mn}}
\newcommand{\mLn}{\mL_{\mn}}
\newcommand{\chibar}{\bar{\chi}}
\newcommand{\ngammadot}{\dot{N}_{\gamma}}
\newcommand{\hi}{H {\footnotesize I}~}
\newcommand{\hii}{H {\footnotesize II}~}

%\newcommand{\mnras}{Mon. Not. R. Astr. Soc.}
%\newcommand{\araa}{Ann. Rev. Astron. Astrophys.}
%\newcommand{\apj}{Ap. J.}
%\newcommand{\apjs}{Ap. J. Supp.}
%\newcommand{\nat}{Nature}
\newcommand{\na}{New Astron.}
%\newcommand{\physrep}{Phys. Reports}


\setlength{\parindent}{0em}
\setlength{\parskip}{2ex}
\textheight 9truein
\textwidth 6.5truein
\addtolength{\oddsidemargin}{+0.25in}
\addtolength{\evensidemargin}{+0.25in}
\addtolength{\topmargin}{+0.5in}

\begin{document}

%\title{Direct Numerical Simulation of Reionization in Large Cosmological Volumes I: Numerical Methods and Tests}
\title{Fully-Coupled Simulation of Cosmic Reionization. I: Numerical Methods and Tests}

%\author{Michael L. Norman, Daniel R. Reynolds, Geoffrey C. So, Robert P. Harkness}

\author{Michael L. Norman\altaffilmark{1,3}}
\author{Daniel R. Reynolds\altaffilmark{4}}
\author{Geoffrey C. So\altaffilmark{1}}
\author{Robert P. Harkness\altaffilmark{2,3}}
\author{John H. Wise\altaffilmark{5}}

\affiliation{
\altaffilmark{1}CASS, University of California, San Diego, 9500 Gilman Drive La Jolla, CA 92093-0424\\
\altaffilmark{2}NICS, Oak Ridge National Laboratory, 1 Bethel Valley Rd, Oak Ridge, TN 37831\\
\altaffilmark{3}SDSC, University of California, San Diego, 9500 Gilman Drive La Jolla, CA 92093-0505\\ 
\altaffilmark{4}Southern Methodist University, 6425 Boaz Ln, Dallas, TX 75205\\
\altaffilmark{5}Center for Relativistic Astrophysics, Georgia Institute of Technology, 837 State St, Atlanta, GA 30332 \\
}



\begin{abstract}
We describe an extension of the {\em Enzo} code to enable \st{the direct numerical} {\bf fully-coupled radiation hydrodynamical} simulation 
of inhomogeneous reionization in large {\bf $\sim (100 Mpc)^3$} cosmological volumes {\bf with thousands to millions of point sources} . \st{By direct we mean} {\bf We solve}
all dynamical, radiative {\bf transfer, thermal, and ionization processes} \st{and chemical properties} \st{are solved} self-consistently on the same mesh, 
as opposed to a postprocessing approach which coarse-grains the radiative transfer. We do,
however, employ a simple subgrid model for star formation which we calibrate to observations. 
The numerical method presented is a modification of an earlier method presented in Reynolds et al., {\bf differing principally in the operator splitting algorithm we we use to advance the system of equations.} Radiation transport is done in the grey flux-limited diffusion (FLD) approximation, which is solved by
implicit time integration split off from the gas energy and ionization equations, which are solved
separately. This results in 
a faster and more robust scheme for cosmological applications compared to the earlier method. 
The FLD equation is solved
using the {\em hypre} optimally scalable geometric multigrid solver from LLNL. 
By treating the ionizing radiation as a grid field as opposed to rays, our method
is scalable with respect to the number of ionizing sources, limited only by the parallel scaling
properties of the radiation solver. We test the speed and accuracy of our approach on a 
number of standard verification and validation tests.  We show {\bf by direct comparison with {\em Enzo}'s adaptive ray tracing method {\em Moray}} that the well-known 
inability of FLD to cast a shadow behind opaque clouds \st{has little effect on the 
photoevaporation timescale of the cloud, or} {\bf has a minor effect on the evolution of ionized volume and mass fractions}
in a reionization simulation validation test. 
We illustrate \st{its application} {\bf an application of our method} to the
problem of inhomogeneous reionization in a \st{20} {\bf80} Mpc comoving box resolved with
\st{$800^3$} {\bf $3200^3$} Eulerian grid cells and dark matter particles. 
\end{abstract}
\keywords{cosmology: theory -- reionization -- methods: numerical -- radiative transfer}

\maketitle


\input{introduction.tex}
\input{mathematical_formulation.tex}
\input{numerical_methods.tex}
\input{tests.tex}
\input{example_simulation.tex}
\input{conclusions.tex}


\bibliography{sources}
\bibliographystyle{apj}
\end{document}
